New Computer Data-Science Set-up: Python (anaconda, pycharm), R (r-studio) and other Programs Installations

### 2019-12-23 nontex/install/windows/fn_installations.Rmd
update installation files

---
title: "New Windows Computer Data-Science Set-up: Python (anaconda, pycharm), R (r-studio) and other Programs Installations"
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

@@

@@

```{bash}
# # if there are bugs
# conda clean --packages
# use conda-forge as main channel, more updated packages
conda config --add channels conda-forge
# normal update
conda update --all
```

@@

@@

@@

conda install -c r r-irkernel
```

@@

1. [download R](https://cloud.r-project.org/)
2. [download R-studio](https://rstudio.com/products/rstudio/download/)
3. Open R-studio and auto-detect R
4. Install packages

note that installation will not work if the machine's version of R is not up to date, this could be the case with default R versions for some linux installations.

```{r}
# Install R-tools
install.packages("installr")
library("installr")
install.Rtools()
# c:\Rtools\bin;
# c:\Rtools\mingw_32\bin;
# c:\Rtools\mingw_64\bin;

# main packagevps
install.packages(c("tidyr", "tidyverse", "tidymodels"))
# development packages
install.packages(c("devtools", "irkernel", "pkgdown", "roxygen2"))
# other packages
install.packages(c("AER", "minpack.lm", "knitr", "kableExtra", "matlab"))

@@

```{bash}
# Set up Channel
conda config --add channels conda-forge
# install key r packages within a r-environment
conda env remove -n r_env
conda create -n r_env r-essentials r-base r-tidyr r-tidyverse r-devtools r-irkernel r-pkgdown r-roxygen2
# see all installed environments
conda env list
# activate an environment to use it, if in base, r does not exist, isolated in r_env
conda activate r_env
# see packages in environment
conda list
# to quit
conda deactivate
```

@@

```{bash}
<!-- for ubuntu -->
wget "https://download1.rstudio.org/desktop/bionic/amd64/rstudio-1.2.5033-amd64.deb"
sudo apt install ./rstudio-1.2.5033-amd64.deb
<!-- for debian -->
wget "https://download1.rstudio.org/desktop/debian9/x86_64/rstudio-1.2.5033-amd64.deb"
sudo apt install ./rstudio-1.2.5033-amd64.deb
conda install -c r rstudio
cd "C:/Program Files/RStudio/bin"
rstudio.exe

@@

@@

# C:\Users\fan\R4Econ\summarize\dist\fst_hist_onevar.Rmd
# C:\Users\fan\R4Econ\support\tibble\fs_tib_basics.Rmd
# C:\Users\fan\R4Econ\optimization\cesloglin\fst_ces_plan_linlog.Rmd
```

@@

## Compile in Atom

Install the [latex](https://github.com/thomasjo/atom-latex) package on atom.

```{bash}
# set TeX path
# C:\texlive\2019\bin\win32
```

@@

- Adobe Acrobat
    + go to adobe website, log in using fwang26@uh.edu
    + go to my account, choose view and download my apps, choose acrobat & PDF, download Acrobat DC
    + UH account can only be activated on two accounts at once, so need to kick other computers out temporarily potentially

>>>>

---
title: "New Computer Data-Science Set-up: Python (anaconda, pycharm), R (r-studio) and other Programs Installations"
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

@@

# Objective

For the following types of programming tasks, install relevant programs, that are cross platform compatible. Test at the end that example programs work.

1. Python
    + Anaconda Dist
    + Pycharm
    + atom hydrogen
2. R
    + R
    + Rstudio
    + atom hydrogen
3. Matlab
    + Matlab
4. Latex and PDF
    + atom latex
    + texstudio
    + texlive
    + okular pdf

Test at the end:

1. Do sample *py* files work?
    + work from command line?
    + work from atom (hydrogen)
2. Do sample *rmd* files work?
    + work from r-studio
    + work from atom (hydrogen), r file
3. Do tex templates compile?
    + tikz files
    + template files with bib

@@

```{bash}
# # if there are bugs
# conda clean --packages
# use conda-forge as main channel, more updated packages
conda config --get channels
conda config --add channels conda-forge
conda config --get channels
# remove conda-forge from channels (do so for main env install)
conda config --remove channels conda-forge
# normal update
conda update --all
```

@@

Fist test python under command line:

@@

## Jupyter and Atom Hydrogen

For both Jupyter as well as Atom, always open from command prompt, open from *Anaconda Prompt* with admin rights.

First, open up *Jupyter Notebook*, does the python kernel work? If does not, uninstall and re-install Anaconda.

```{bash}
# start jupyter lab, from prompt
jupyter lab
```

Now test python within atom:

1. install hydrogen in atom. Follow [instructions](http://nteract.io/atom).
    - terminal in windows or linux after installation: *apm install hydrogen*
    - apm is the atom package manager.
2. go to the programs above and try to run.

```{bash}
# Once atom is installed, can use apm to install packages
apm install hydrogen
# Inside Conda open up atom
atom
# open up any py from from within, try 1+1, does it work?
```

@@

# conda install -c r r-irkernel
```

@@

1. [download R](https://cloud.r-project.org/)
    + for debian: [Johannes Ranke](https://cran.r-project.org/bin/linux/debian/)
2. [download R-studio](https://rstudio.com/products/rstudio/download/)
3. Open R-studio and auto-detect R
4. Install packages

note that installation will not work if the machine's version of R is not up to date, this could be the case with default R versions for some linux installations. For Debian for example, need to update the *source.list* to include sources that have more recent versions of R.

Additionally, need to install *libcurl4-openssl-dev* and *libssl-dev* from command prompt first inside Debian.

```{bash}
# Go to get latesdebian latest r sources.list
cat /etc/apt/sources.list
# Install this First (should already be installed)
sudo apt install dirmngr

# Debian R is maintained by Johannes Ranke, copied from https://cran.r-project.org/bin/linux/debian/:
apt-key adv --keyserver keys.gnupg.net --recv-key 'E19F5F87128899B192B1A2C2AD5F960A256A04AF'
# Add to source.list, for debian stretch (9)
# sudo su added for security issue as super-user
sudo su -c "sudo echo 'deb http://cloud.r-project.org/bin/linux/debian stretch-cran35/' >> /etc/apt/sources.list"
# if added wrong lines, delete 3rd line
sudo sed '3d' /etc/apt/sources.list

# Update and Install R, should say updated from cloud.r
sudo apt-get update
sudo apt-get install r-base r-base-dev

# Also install these, otherwise r-packages do not install:
sudo apt-get install libcurl4-openssl-dev
sudo apt-get install libssl-dev
```

```{r}
# Install R-tools, RTOOLS is for WINDOWS
install.packages("installr")
library("installr")
install.Rtools()
# c:/Rtools/bin;
# c:/Rtools/mingw_32/bin;
# c:/Rtools/mingw_64/bin;

# main packagevps
# tidymodel installation on linux difficult
install.packages(c("tidyr", "tidyverse", "tidymodels"))
# development packages
install.packages(c("devtools", "irkernel", "pkgdown", "roxygen2"))
# other packages
install.packages(c("AER", "minpack.lm", "knitr", "kableExtra", "matlab"))

# Install my package
devtools::install_github("fanwangecon/R4Econ")

# Kernel
install.packages(c('repr', 'IRdisplay', 'evaluate', 'crayon', 'pbdZMQ', 'devtools', 'uuid', 'digest'))
devtools::install_github('IRkernel/IRkernel')
IRkernel::installspec()
```

@@

```{bash}
# Set up Channel, for R, to get latest r-packages
# do not set-up channels to prioritize conda-forge for main env.
conda config --add channels conda-forge
# install key r packages within a r-environment
conda env remove -n r_env
conda create -n r_env r-essentials r-base r-tidyr r-tidyverse r-devtools r-irkernel r-pkgdown r-roxygen2
# see all installed environments
conda env list
# activate an environment to use it, if in base, r does not exist, isolated in r_env
conda activate r_env
# see packages in environment
conda list
# to quit
conda deactivate
```

@@

```{bash}
<!-- for ubuntu -->
wget "https://download1.rstudio.org/desktop/bionic/amd64/rstudio-1.2.5033-amd64.deb"
<!-- for debian -->
sudo apt install ./rstudio-1.2.5033-amd64.deb
wget "https://download1.rstudio.org/desktop/debian9/x86_64/rstudio-1.2.5033-amd64.deb"
sudo apt install ./rstudio-1.2.5033-amd64.deb
conda install -c r rstudio
cd "C:/Program Files/RStudio/bin"
rstudio.exe

@@

# for r-kernel to work in Atom
install.packages(c('repr', 'IRdisplay', 'evaluate', 'crayon', 'pbdZMQ', 'devtools', 'uuid', 'digest'))
devtools::install_github('IRkernel/IRkernel')
IRkernel::installspec()

@@

# C:/Users/fan/R4Econ/summarize/dist/fst_hist_onevar.Rmd
# C:/Users/fan/R4Econ/support/tibble/fs_tib_basics.Rmd
# C:/Users/fan/R4Econ/optimization/cesloglin/fst_ces_plan_linlog.Rmd
```

@@

## Compile in Atom and Texstudio

Install the following two packages in atom:

1. [latex](https://github.com/thomasjo/atom-latex) package on atom.
2. [language-latex](https://atom.io/packages/language-latex) package on atom

In Atom Latex setting, note that on different platforms, paths to texlive are different:

- Path:
    + Windows: *C:/texlive/2019/bin/win32*
    + Chromebook Linux: */usr/bin/latex*
- Which PDF viewer to use:
    + download Okular and use [Okular](https://okular.kde.org/)

```{bash}
# set TeX path
# C:/texlive/2019/bin/win32

# Latex, after installation
# see where texlive appears under root folder directories
sudo find / -name "texlive"
# see latex version, if texlive is used
latex -v
# see main exe directory
which latex
# /usr/bin/latex
```

Go to Tex4Econ and compile various template files and other types of files there to see if tex is working.

Compile the same files in texstudio.

@@

**PDF**

- PDF Okular
    + Linux: *sudo apt-get install okular*
    + Windows: install okular from windows app store
- PDF Adobe Acrobat
    + go to adobe website, log in using fwang26@uh.edu
    + go to my account, choose view and download my apps, choose acrobat & PDF, download Acrobat DC
    + UH account can only be activated on two accounts at once, so need to kick other computers out temporarily potentially

**VPN**

- [Proton VPN](https://protonvpn.com/): unlimited usage free version available.
